\vspace{-2cm} Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States

### 2019-07-31 paper/manuscript_psrmNote.tex
comment on the longevity of the software we use (to address Gill's comment)

\noindent PA Letter Requirements:\\
2-4 pages\\
no longer than 1500-3000 words\\
1-3 small display items (figures, tables, or equations)\\
200-300 word abstract\\

@@

\section{The Web to Text Pipeline}
In this section, we describe our methodological pipeline, with which we take an archive of website files, and output a corpus of formatted plain text documents. We address three methodological challenges.  First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step is aimed at extracting clean plain text from this heterogeneous file base. The second step in our pipeline is to process the text to remove language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments. Finally, these tools need to work consistently across all of the websites in our corpus, in spite of the fact that relevant information is stored and structured in different ways.%Some of the steps we take in this processing pipeline are universally applicable in the analysis of textual data, and some of them are most appropriate for the particular type of text analysis that we apply to this data---statistical topic modeling. We will clarify this distinction as we describe steps in our pipeline.

@@

\subsubsection{Extracting Text from HTML}
The HTML files that websites are comprised of contain a large amount of useful information, but also completely irrelevant text such as menus, navigational elements and other boilerplate. The side-by-side screenshots presented in Figure \ref{fig:garysbs} convey the challenges presented by extracting content for text analysis for websites. The textual content that is substantive and unique to the Gary, IN homepage is the Mayor's message depicted in Figure \ref{garymayor}. The top row of Figure \ref{fig:garysbs} presents the complete homepage, along with all of the text that can be naively extracted from the site. The Mayor's message represents a relatively small fraction of the total text on the page.

@@

\subsubsection{Extracting Text from PDF, XML, DOC, DOCX and TXT}
Other files are read in through the \texttt{readtext} R package \citep{readtext}, which is a wrapper for a set of parsers.\footnote{\texttt{readtext} determines a document's type solely through its ending---so the conversion described above is necessary.}$^{,}$\footnote{\texttt{readtext} also contains an HTML parser, but it does not eliminate boilerplate like boilerpipe.} The breakdown of all files by type is given in the online appendix. The most frequent file type besides HTML is PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

To study content differences between government websites based on mayoral partisanship, we draw upon a recently-developed class model for text, the structural topic model (STM), developed by \citet{Roberts2014}. Building on the conception of ``topics'' in Latent Dirichlet Allocation, in the STM a topic is a multinomial distribution defined on the word types in the corpus dictionary. The log-odds of the topic probabilities in each document-specific multinomial distribution over topics are drawn from a multivariate normal distribution in which the topic-specific means are determined by a linear regression function that associates document-attributed covariates with topics. For example, in the context of municipal website content, the structural topic model can be used to estimate a regression coefficient that defines the linear relationship between the log-odds of the municipality's population and the log-odds of each topic. For our primary empirical investigation, the STM provides with a tool with which to estimate the relationship between the party of the city's mayor and the prevalence of each topic we estimate. We also include the municipality's population and median income as covariates. Further details on and results from our STM specification can be found in the online appendix.

@@

The results are shown in Table \ref{tabSTMtopwords60}. First, it is notable that the 95\% credible includes zero in only seven of the sixty topics, indicating that the topics discussed on city websites varies systematically with the partisanship of the mayor. Many of the topics associated with Democrats fit with what we understand to be national party priorities. Topic \textbf{52}, on affordable housing, clearly resonates with the Democratic party's appeal to low-income voters. Topic {\bf 6} ('race', 'islander','census,'female') covers racial and gender identity issues. Similarly, employee rights and benefits are represented in topics \textbf{10} and \textbf{29}. Democrats also exhibit a strong preference for words related to public finances, such as Topic \textbf{58} ('budget', 'revenue', 'expenditure'), Topic {\bf 45} ('asset', 'actuarial', 'liability', 'financial'), Topic \textbf{35} ('bond', 'obligation', 'proceeds') as well as Topic \textbf{55} ('taxable', 'deed', 'value'). We suspect that the association of Democratic mayors with finance-related terms is indicative of a greater willingness to emphasize the city's efforts to raise and spend money, and take credit for those efforts (e.g., the Gary, IN example in Figure \ref{fig:garymayor}). This finding is consistent with \citep{Einstein2015}, who show that Democratic mayors tend to favor greater spending. A second, consistent Democratic focus appears to be law enforcement: The most Democratic topic, \textbf{59} ('burglary', 'robbery', 'theft', 'homicide') is clearly focused on crime. On the one hand, Democratic partisans have a more negative perception of the police, rating it considerably more negatively on the appropriate use of force and the equal treatment of minorities \citep{Brown2017}. On the other hand, the literature has also shown that cities with a higher Democratic vote share spend more on law enforcement, even after controlling for crime \citep{Einstein2015}.

City websites with Republican mayors, meanwhile, exhibit a pronounced focus on the essential functions of government. Basic utilities such as energy (Topic \textbf{20}), fire protection (Topic \textbf{51}), vaccination (Topic \textbf{2}),  and sanitation (Topic \textbf{47}), are prevalent among cities with Republican mayors. These basic service topics cannot be found among topics prevalent in cities with Democratic mayors. Similarly, zoning issues figure prominently in the set of republican topic (Topic \textbf{19}), which fits with the findings of \citet{sorens2018effects} that Republicans are more supportive of restrictive residential zoning rules. The Democratic topics also include one that is somewhat focused on zoning Topic {\bf 39} ('downtown', 'mixed', 'density'), but emphasizes mixed-use zoning---a loosening of conventional single-use zoning rules.

@@

The biggest limitation in our pipeline, and an open area for future research, is the reliance on \texttt{wget} to gather the initial website files. By using \texttt{wget}, we miss content that is displayed dynamically on websites using JavaScript. For any one website, it would be possible to customize a routine with \texttt{Selenium} to access dynamic elements, but the process would need to be customized for each website. \footnote{ We investigated whether the presence of JavaScript was related to the amount of text we gathered from the website. We calculated the correlation between the number of <script> HTML tags on a city's website, which indicate the use of JavaScript on a site, and the number of text tokens we scrape from the site. This correlation is -0.059, which indicates a very weak relationship between the use of JavaScript and the amount of text scraped from the site.}

>>>>

@@

\section{The Web to Text Pipeline}
In this section, we describe our methodological pipeline, with which we take an archive of website files, and output a corpus of formatted plain text documents. We address three methodological challenges.  First, though they contain significant amounts of text, websites are not comprised of clean plain text files. Rather, the files available at websites are of multiple types, including HTML, PDF, word processor, plain text, and image files. The first step is aimed at extracting clean plain text from this heterogeneous file base. The second step in our pipeline is to process the text to remove language that is effective at differentiating one website from another but is uninformative regarding policy or political differences between governments. Finally, these tools need to work consistently across all of the websites in our corpus, in spite of the fact that relevant information is stored and structured in different ways. We make a software recommendation for each of these steps and gather most of them our R package, \texttt{gov2text}. All of the recommended software is either well-established in the natural language processing community, or part of the Unix ecosystem. As such, all of it is free, open source, well-developed and will continue to be supported by a dedicated community.%Some of the steps we take in this processing pipeline are universally applicable in the analysis of textual data, and some of them are most appropriate for the particular type of text analysis that we apply to this data---statistical topic modeling. We will clarify this distinction as we describe steps in our pipeline.

@@

\subsubsection{Extracting Text from HTML}
The HTML files that websites are comprised of contain a large amount of useful information, but also completely irrelevant text such as menus, navigational elements and other boilerplate. The side-by-side screenshots presented in Figure \ref{fig:garysbs} convey the challenges presented by extracting content for text analysis for websites. The textual content that is substantive and unique to the Gary, IN homepage is the Mayor's message depicted in Figure \ref{fig:garymayor}. The top row of Figure \ref{fig:garysbs} presents the complete homepage, along with all of the text that can be naively extracted from the site. The Mayor's message represents a relatively small fraction of the total text on the page.

@@

\subsubsection{Extracting Text from PDF, DOC, DOCX and TXT}
Other files are read in through the \texttt{readtext} R package \citep{readtext}, which is a wrapper for a set of parsers.\footnote{\texttt{readtext} determines a document's type solely through its ending---so the conversion described above is necessary.}$^{,}$\footnote{\texttt{readtext} also contains an HTML parser, but it does not eliminate boilerplate like boilerpipe.} The breakdown of all files by type is given in the online appendix. The most frequent file type besides HTML is PDF, from which we are able to extract a substantial amount of usable text. Files of type DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

To study content differences between government websites based on mayoral partisanship, we draw upon a recently-developed class model for text, the structural topic model (STM), developed by \citet{Roberts2014}. Building on the conception of ``topics'' in Latent Dirichlet Allocation, in the STM a topic is a multinomial distribution defined on the word types in the corpus dictionary. The log-odds of the topic probabilities in each document-specific multinomial distribution over topics are drawn from a multivariate normal distribution in which the topic-specific means are determined by a linear regression function that associates document-attributed covariates with topics. For example, in the context of municipal website content, the structural topic model can be used to estimate a regression coefficient that defines the linear relationship between the log-odds of the municipality's population and the log-odds of each topic. For our primary empirical investigation, the STM provides a tool to estimate the relationship between the party of the city's mayor and the prevalence of each topic. We also include the municipality's population and median income as covariates. Further details on and results from our STM specification can be found in the online appendix.

@@

The results are shown in Table \ref{tabSTMtopwords60}. First, it is notable that the 95\% credible interval includes zero in only seven of the sixty topics, indicating that the topics discussed on city websites varies systematically with the partisanship of the mayor. Many of the topics associated with Democrats fit with what we understand to be national party priorities. Topic \textbf{52}, on affordable housing, clearly resonates with the Democratic party's appeal to low-income voters. Topic {\bf 6} ('race', 'islander', 'census, 'female') covers racial and gender identity issues. Similarly, employee rights and benefits are represented in topics \textbf{10} and \textbf{29}. Democrats also exhibit a strong preference for words related to public finances, such as Topic \textbf{58} ('budget', 'revenue', 'expenditure'), Topic {\bf 45} ('asset', 'actuarial', 'liability', 'financial'), Topic \textbf{35} ('bond', 'obligation', 'proceeds') as well as Topic \textbf{55} ('taxable', 'deed', 'value'). We suspect that the association of Democratic mayors with finance-related terms is indicative of a greater willingness to emphasize the city's efforts to raise and spend money, and take credit for those efforts (e.g., the Gary, IN example in Figure \ref{fig:garymayor}). This finding is consistent with \cite{Einstein2015}, who show that Democratic mayors tend to favor greater spending. A second, consistent Democratic focus appears to be law enforcement: The most Democratic topic, \textbf{59} ('burglary', 'robbery', 'theft', 'homicide') is clearly focused on crime. On the one hand, Democratic partisans have a more negative perception of the police, rating it considerably more negatively on the appropriate use of force and the equal treatment of minorities \citep{Brown2017}. On the other hand, the literature has also shown that cities with a higher Democratic vote share spend more on law enforcement, even after controlling for crime \citep{Einstein2015}.

City websites with Republican mayors, meanwhile, exhibit a pronounced focus on the essential functions of government. Basic utilities such as energy (Topic \textbf{20}), fire protection (Topic \textbf{51}), vaccination (Topic \textbf{2}),  and sanitation (Topic \textbf{47}), are prevalent among cities with Republican mayors. These basic service topics cannot be found among topics prevalent in cities with Democratic mayors. Similarly, zoning issues figure prominently in the set of republican topic (Topic \textbf{19}), which fits with the findings of \citet{sorens2018effects} that Republicans are more supportive of restrictive residential zoning rules. The Democratic topics also include one that is somewhat focused on zoning, Topic {\bf 39} ('downtown', 'mixed', 'density'), but emphasizes mixed-use zoning---a loosening of conventional single-use zoning rules.

@@

The biggest limitation in our pipeline, and an open area for future research, is the reliance on \texttt{wget} to gather the initial website files. By using \texttt{wget}, we miss content that is displayed dynamically on websites using JavaScript. For any one website, it would be possible to customize a routine with \texttt{Selenium} to access dynamic elements, but the process would need to be customized for each website.\footnote{ We investigated whether the presence of JavaScript was related to the amount of text we gathered from the website. We calculated the correlation between the number of <script> HTML tags on a city's website, which indicate the use of JavaScript on a site, and the number of text tokens we scrape from the site. This correlation is -0.059, which indicates a very weak relationship between the use of JavaScript and the amount of text scraped from the site.}
